---
title: "Project1 "
author: "Justin Herman"
date: "8/29/2019"
output:
  html_document:
    theme: "simplex"
    highlight: 'pygments'
---

<style>h1{text-transform: capitalize}</style>
<style>h2{text-transform: capitalize}</style>
<style>h3{text-transform: capitalize}</style>
<style>p:first-letter {text-transform: capitalize}</style>
<style>li:first-letter {text-transform: capitalize}</style>

<style type="text/css">body{ /* Normal  */ font-size: 18px;}}</style>



# Overview {.tabset .tabset-fade}

```{r,echo=F}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```



+ Extra things to look at psych describe package has standard deviation skew and kurtosis
+ Hmisc has corr to look for colinearity 

```{r,echo=FALSE,warning=F,message=F}
library(psych)
library(corrplot)
library(Hmisc)
library(usmap)
library(ggplot2)
library(ggplot2)
library(sp)
library(tidyverse)
library(ggmap)
library(scales)
##
psych::describe(inc)

cor(inc[,c(3,4,6)])


```

## Questions  {.tabset .tabset-fade}

+ 12 values are missing in employee numbers, as directions want us to ignore those observations, I will drop them 

### Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

#### First graph uses color graident to show total companies by state 

```{r,echo=F,warning=F,message=F}

##data transformations
inc <- inc %>% 
    mutate(state=as.character(State),
           city=as.character(City))

## drop 13 observations which are missing employment data
inc <- inc[complete.cases(inc), ]

## build dataframe to use with mapping that coutns employment by state
map_df <- inc %>% 
    group_by(state) %>% 
    summarise(count_per_state=length(Name)) %>% 
    ungroup() 

##plot employment with graident color from white to red

plot_usmap(data = map_df, values = 'count_per_state', lines = "red") + 
  scale_fill_continuous(
    low = "white", high = "red", name = "count_per_state", label = scales::comma
  ) + theme(legend.position = "right")
```


#### Next I import google maps api log and lat data based on city data in dataframe and creates a ggpoint plot of all companies in america


```{r,echo=F, warning=F,message=F }
## use google api to convert cities into lat/long
## Due to costs I load in csv from query call

# api <- readLines("E:\\Google_api.txt")
# register_google(key = api)

# Load in dataset 
locations_df2 <-read_csv("geo_locations.csv")    
locations_df2 <- locations_df2 %>% 
    select(geo_state=state,city=City,lon,lat)

joined_locations <- plyr::join(inc,locations_df2,'city')

### Something went wrong with join, we have over 100k rows
### first step, drop duplicates
joined_locations <- unique(joined_locations)
```

Quick data check

```{r,echo=F}
## now were down to 5900, lets see two " nearly identical rows"
joined_locations[4,]==joined_locations[5,]

## get rid of mismatches, google returned multiple when city could be in multiple states.  
## Easy fix I just said drop any row where google state is different from dataset state
joined_locations <- joined_locations %>% 
    filter(geo_state==state)

## print dim
dim(joined_locations)


##plot by location with size based on number employees
map<-get_map(location='united states', zoom=4, maptype = "terrain",
             source='google',color='color')

ggmap(map) + geom_point(
        aes(x=lon, y=lat, show_guide = TRUE), 
        data=joined_locations, alpha=.5, na.rm = T)  + 
        scale_color_gradient(low="beige", high="blue")

```


#### Next Approach uses simple bar plot

```{r,echo=F}


## bar plot
ggplot(data=map_df, aes(x=reorder(state,count_per_state), y=count_per_state)) +
  geom_bar(stat="identity")+
    theme(text = element_text(size=8))+
    coord_flip()
```



#### Next two approaches use dotplot and lollipop printed in reverse orders

```{r,echo=F}
ggplot(map_df, aes(x=reorder(state,count_per_state), y=count_per_state, label=count_per_state)) + 
  geom_point(stat='identity', aes(), size=6)  +
  scale_color_manual(name="Mileage", 
                     labels = c("Above Average", "Below Average"), 
                     values = c("above"="#00ba38", "below"="#f8766d")) + 
  geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot", 
       subtitle="Total Company by State") + 
 # ylim(-2.5, 2.5) +
    theme(text = element_text(size=8))+
  coord_flip()
 

## 
ggplot(map_df, aes(x=reorder(state,-count_per_state), y=count_per_state, label=count_per_state)) + 
  geom_point(stat='identity', size=8)  +
  geom_segment(aes(y = 0, 
                   x = state, 
                   yend = count_per_state, 
                   xend = state), 
               color = "blue") +
  geom_text(color="white", size=2) +
  labs(title="Diverging Lollipop Chart", 
       subtitle="Total Company by State") + 
    theme(text = element_text(size=8))+
  coord_flip()



```


### Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.


#### Which State?

+ As seen below, NY is the state we will focus on

```{r,echo=F}
## sort df with company count by state and identify state for research
##NY
map_df <- map_df %>% 
    arrange(-count_per_state)
map_df[3,]
```


Below I print several visualizations.

```{r,echo=F}
NY_df <- inc %>% 
    filter(state=="NY") %>% 
    mutate(Industry=as.character(Industry)) %>% 
    filter()

##option 1 without dots
g <- ggplot(NY_df, aes(Industry, Employees))
g + geom_boxplot() + 
  # geom_dotplot(binaxis='y', 
  #              stackdir='center', 
  #              dotsize = .5, 
  #              fill="red") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    scale_y_continuous(trans = 'log10',
                        breaks = trans_breaks('log10', function(x) 10^x),
                        labels = trans_format('log10', math_format(10^.x)))
  labs(title="NYS Employment by Industry",
       x="Industry",
       y="Log Transformed Employment by company")+
  coord_flip()


g <- ggplot(NY_df, aes(Industry,Employees))
g + geom_boxplot() + 
  geom_dotplot(binaxis='y',
               stackdir='center',
               dotsize = .4,
               fill="red") +
    scale_y_continuous(trans = 'log10',
                        breaks = trans_breaks('log10', function(x) 10^x),
                        labels = trans_format('log10', math_format(10^.x)))+
  theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
  labs(title="NYS Employment by Industry",
       x="Industry",
       y="Log Transformed Employment by company")


g <- ggplot(NY_df, aes(Industry, log10(Employees)))
g +  geom_violin()+
  theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
  labs(title="NYS Employment by Industry",
       x="Industry",
       y="Log Transformed Employment by company")

```

### Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.



```{r}



industry_revenue <- inc %>% 
    mutate(rev_per_employee= Revenue/Employees) 

    

g <- ggplot(industry_revenue, aes(Industry, rev_per_employee,10))
g + geom_boxplot() + 
  # geom_dotplot(binaxis='y', 
  #              stackdir='center', 
  #              dotsize = .5, 
  #              fill="red") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
  labs(title="US Revenue per Employee by industry",
       x="Industry",
       y="Log Transformed Revenue per Employee by company")+
    scale_y_continuous(trans = 'log10',
                        breaks = trans_breaks('log10', function(x) 10^x),
                        labels = trans_format('log10', math_format(10^.x)))+
  coord_flip()
```
